---
title: "Exploratory Visualisations"
author: "Aditya Khan"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

```{r echo = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(data.table)
library(ggplot2)
library(leaflet)
library(readxl)
library(geosphere)
library(knitr)
library(kableExtra)
library(tidytext)
library(wordcloud)
library(tm)
library(RColorBrewer)  
library(gridExtra)
library(htmltools)
library(cowplot)
library(pastecs)
library(rpart)
library(rpart.plot)
library(randomForest)
library(gbm)
library(xgboost)
library(caret)
#install.packages("car")
library(car)
library(faraway)
library(lme4)
library(nlme)
library(lattice)
library(mlmhelpr)
library(plotly)
library(widgetframe)
library(nlme)
library(googledrive)
```

**Important Note:** On this page, please note that all the visualisations that are on the report, are here as well. But important details about data collection are ommitted for brevity. It is probably better to read the full report first, as a result. To follow the guidelines, there are at most *five* interactive visualisations (there are more non-interactive ones though). 

# Data Collection 

All of the datasets were downloaded from the websites mentioned in the references sections. In the references section, [This](http://insideairbnb.com/get-the-data/) is the website to download the AirBnB data. Navigate to the section titled Toronto, and download "listings.csv.gz", under February 14, 2024. [This](https://data.torontopolice.on.ca/datasets/TorontoPS::major-crime-indicators-open-data/about) is the website to download the MCI Open Data. [This](https://open.toronto.ca/dataset/neighbourhood-profiles/ ) is the website to download the Neighbourhood Profile data. For the latter two, there is a download button on the website. The former two datasets are in CSV format. The final dataset mentioned, is in XLSX format. 

Note that the AirBnB data and the income data can be found on the Github repo, but the crime data cannot, due to the size of the dataset. Instructions for how to obtain it, are given in the Readme in the data folder. 

```{r echo = FALSE, message=FALSE, warning=FALSE, results='hide'}
abnb <- read.csv("data/feb 24.csv", na.strings = c("", NA))
setDT(abnb)

crime <- read.csv("data/Major_Crime_Indicators_Open_Data.csv")
setDT(crime)

nbd_profiles <- read_xlsx("data/neighbourhood-profiles-2021-158-model.xlsx")

```

```{r echo = FALSE, message=FALSE, warning=FALSE, results='hide'}

false_x <- crime[crime$LONG_WGS84 == 0][1]$X
false_y <- crime[crime$LONG_WGS84 == 0][1]$Y
crime[] <- lapply(crime, function(x) gsub("NSA", NA, x))

crime$LONG_WGS84[crime$LONG_WGS84 == 0] <- NA
crime$LAT_WGS84[crime$LAT_WGS84 == 0] <- NA
crime$X[crime$X == false_x] <- NA
crime$Y[crime$Y == false_y] <- NA

crime_sub <- crime[, c("OCC_YEAR", "OCC_MONTH", "OCC_DAY", "NEIGHBOURHOOD_158", "LONG_WGS84", "LAT_WGS84")]
crime_sub <- na.omit(crime_sub)
crime_sub$occ_month_int <- match(crime_sub$OCC_MONTH, month.name)
crime_sub$DATE_OCC <- make_datetime(year = as.numeric(crime_sub$OCC_YEAR), month = crime_sub$occ_month_int, day = as.numeric(crime_sub$OCC_DAY))
target_date <- as.Date("2024-02-14")
crime_sub <- filter(crime_sub, DATE_OCC >= (target_date - 365), DATE_OCC <= target_date)
crime_count <- crime_sub[, .(crime_count = .N), by = NEIGHBOURHOOD_158]
crime_count$crime_count <- crime_count$crime_count / 365

median_data <- subset(nbd_profiles, `Neighbourhood Name` == "Median total income of household in 2020 ($)")

med_t <- t(median_data[-1])
med_dt <- setDT(data.frame(nbd = rownames(med_t), nbd.med_hh = as.numeric(med_t)))

abnb[] <- lapply(abnb, function(x) gsub("N/A", NA, x))
convert_price_to_double <- function(price_char) {
  price_double <- as.numeric(sub("\\$", "", price_char))
  return(price_double)
}

abnb$price <- sapply(abnb$price, convert_price_to_double)
abnb_sub <- abnb[, c("neighborhood_overview", "neighbourhood_cleansed", "latitude", "longitude", "property_type", "price")]
abnb_sub$latitude <- as.numeric(abnb_sub$latitude)
abnb_sub$longitude <- as.numeric(abnb_sub$longitude)

abnb_sub <- abnb_sub[complete.cases(abnb_sub$price), ]
abnb_sub <- abnb_sub[abnb_sub$price > 1]

abnb_sub$property_type <- ifelse(grepl("^Entire", abnb_sub$property_type), "Entire property",
                                 ifelse(grepl("^Private room", abnb_sub$property_type), "Private room",
                                        ifelse(grepl("^Shared room", abnb_sub$property_type), "Shared room",
                                               ifelse(grepl("^Room", abnb_sub$property_type), "Unspecified room", abnb_sub$property_type))))

merged <- merge(
  x     = abnb_sub,      
  y     = crime_count, 
  by.x  = c("neighbourhood_cleansed"),
  by.y  = c("NEIGHBOURHOOD_158"), 
  all.x = FALSE,      
  all.y = FALSE
)

merged <- merge(
  x     = merged,      
  y     = med_dt, 
  by.x  = c("neighbourhood_cleansed"),
  by.y  = c("nbd"), 
  all.x = FALSE,      
  all.y = FALSE
)

new <- c("nbd", "nbd_desc", "lat", "lon", "property_type", "price", "nbd.crime_count", "nbd.med_hh")
old <- colnames(merged)
setnames(merged, old, new)

downtown_coords <- c(43.6515, -79.3835)
merged$distance_from_downtown <- distHaversine(downtown_coords, merged[, c("lat", "lon")]) / 1000 

merged$nbd.med_hh_scaled <- merged$nbd.med_hh / 100

merged$property_type <- as.factor(merged$property_type)
merged$nbd <- as.factor(merged$nbd)

property_table <- table(merged$property_type)

remove_type <- c("Boat", "Camper/RV", "Casa particular", "Castle", "Tiny home", "Treehouse", "Unspecified room")

merged <- merged[!(property_type %in% remove_type)]

merged <- merged[, -c("nbd.med_hh")]

```



```{r echo = FALSE, message=FALSE, warning=FALSE}

variables <- c("nbd", "nbd_desc", "lat", "lon", "property_type", 
               "price", "nbd.crime_count", "distance_from_downtown", "nbd.med_hh_scaled")

variable_description <- c("Neighborhood", "Neighbourhood Description Given by Host", 
                           "Latitude", "Longitude", "Property Type", 
                           "Price of Property", "Crime Count in Neighborhood (/day) in Past Year", 
                           "Distance from Downtown (km)",
                           "Scaled Median Household Income in Neighborhood (/$100)")

variable_type <- c("Categorical", "Not Used as Covariate", "Numerical Predictor", "Numerical Predictor", "Categorical Confounder", 
                   "Numerical Response", "Numerical Predictor", "Numerical Predictor", 
                   "Numerical Response")

variable_table <- data.frame(Variable = variables,
                              Meaning = variable_description,
                              Role = variable_type)

kable(variable_table) %>%
  kable_styling(full_width = FALSE, "striped") %>%
  add_header_above(c("Variable" = 1, "Information" = 2)) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, bold = TRUE) 
```

**Table 1: Variable Descriptions**

After merging and cleaning the three datasets described above, we end up with eleven columns, and 8982 observations. We specify the variables we will be using in the rest of the analysis and what they represent in table 1 above. Following this, we proceed to conduct some exploratory analysis on our data. 

# Visualisations

## Analysing Text Data

Now we do We start by analysing the neighbourhood descriptions that hosts write on the listing page, detailing qualities of the neighbourhood. Particularly, we want to see if the words they use differ by price level. Here, we defining a price to be "low" (if between 0-quartile 1), "medium" (if between quartile 1-3), or "high" (above quartile 3). 

```{r echo = FALSE, message=FALSE, warning=FALSE}

prices <- merged$price
p.q <- quantile(prices, na.rm = TRUE)
merged$price_level <- ifelse(prices >=  0 & prices <= p.q[2], "low",
                                ifelse(prices > p.q[2] & prices <= p.q[4], "medium",
                                       ifelse(prices > p.q[4], "high", NA)))


incomes <- merged$nbd.med_hh_scaled
hh.q <- quantile(incomes, na.rm = TRUE)
merged$income_level <- ifelse(incomes >=  0 & incomes <= hh.q[2], "low",
                              ifelse(incomes > hh.q[2] & incomes <= hh.q[4], "medium",
                                     ifelse(incomes > hh.q[4], "high", NA)))

cc <- merged$nbd.crime_count
cc.q <- quantile(cc, na.rm = TRUE)
merged$cc_level <- ifelse(cc >=  0 & cc <= cc.q[2], "low",
                          ifelse(cc > cc.q[2] & cc <= cc.q[4], "medium",
                                 ifelse(cc > cc.q[4], "high", NA)))

merged$income_level <- as.factor(merged$income_level)
merged$cc_level <- as.factor(merged$cc_level)


```


```{r fig.width=7, fig.height=7, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center'}
sw <- c(stopwords("english"), "br", "toronto", "neighbourhood", "neighborhood")
merged.text <- merged[!is.na(merged$nbd_desc), ]
merged.text_low <- merged.text[price_level == "low"]
merged.text_mid <- merged.text[price_level == "medium"]
merged.text_high <- merged.text[price_level == "high"]

palette_low <- brewer.pal(9, "Blues")[4:9]
palette_mid <- brewer.pal(9, "Purples")[4:9]
palette_high <- brewer.pal(9, "Oranges")[4:9]

tokens_low <- merged.text_low |>
  select(nbd_desc) |>
  unnest_tokens(token, nbd_desc) |>
  filter(!token %in% sw) |>
  filter(!str_detect(token, "[[:digit:]]+")) |>
  count(token, sort = TRUE) |>
  head(20)

tokens_mid <- merged.text_mid |>
  select(nbd_desc) |>
  unnest_tokens(token, nbd_desc) |>
  filter(!token %in% sw) |>
  filter(!str_detect(token, "[[:digit:]]+")) |>
  count(token, sort = TRUE) |>
  head(20)

tokens_high <- merged.text_high |>
  select(nbd_desc) |>
  unnest_tokens(token, nbd_desc) |>
  filter(!token %in% sw) |>
  filter(!str_detect(token, "[[:digit:]]+")) |>
  count(token, sort = TRUE) |>
  head(20)

layout(matrix(1:3, nrow = 1)) 

par(mar = c(2, 2, 2, 2)) 
wordcloud(words = tokens_low$token, 
          freq = tokens_low$n,
          scale = c(2.5, 0.25),  
          max.words = 50,
          colors = palette_low)     
text(x = 0.5, y = -0.2, labels = "Low Price", cex = 1.5, adj = 0.5)

wordcloud(words = tokens_mid$token, 
          freq = tokens_mid$n,
          scale = c(2.5, 0.25),
          max.words = 50,
          colors = palette_mid)
text(x = 0.5, y = -0.2, labels = "Mid Price", cex = 1.5, adj = 0.5)

wordcloud(words = tokens_high$token, 
          freq = tokens_high$n,
          scale = c(2.5, 0.25),
          max.words = 50,
          colors = palette_high)
text(x = 0.5, y = -0.2, labels = "High Price", cex = 1.5, adj = 0.5)


```


**Figure 1: Word Clouds of Neighbourhood Descriptions, by Price Level**

We see in Figure 1 that the words used are almost the same across all three price levels. Namely, there is an emphasis on walkability, restaurants and shops, and being close to parks. On one hand, this may just be telling us that hosts like to emphasise those facts irrespective of price levels since they know they will appeal to the consumer. 

On the other hand, it is notable that all of the things that are emphasised in the word clouds are things that characterise downtown: namely quick access to amenities and easy walkability. If we interpret these words as what hosts think will appeal to a customer, clearly that means the customer demands these things. And if they're found in downtown, then we would expect demand for AirBnBs near downtown to be higher - and hence, raise the cost of the listing. So this can suggest in a weak sense that proximity to downtown may have some influence on price (in fact, this is verified explicitly later in the map visualisations on the website). 

One thing that this text analysis does show, is that specific amenities could potentially be predictive of price. Namely, restaurants, shops and parks. These feature prominently in the wordclouds. Hence, we create binary variables near.rest, near.shop, near.park for those amenities respectively, evaluating to 1 if the term shows up in the neighbourhood description and 0 otherwise. 

Beyond what we did above, we can also do a more granular analysis, by additionally seeing if the word clouds change, as we consider only listings of a certain property type. In our case, subsetting our data only for listings that yield the "entire property" or a "private room" yield nearly identical wordclouds to that in figure 1. 

```{r fig.width=7, fig.height=7, echo = FALSE, message=FALSE, warning=FALSE, fig.align='center'}

merged.sr <- merged[property_type == "Shared room"]
sw <- c(stopwords("english"), "br", "toronto", "neighbourhood", "neighborhood")
merged.text <- merged.sr[!is.na(merged.sr$nbd_desc), ]
merged.text_low <- merged.text[price_level == "low"]
merged.text_mid <- merged.text[price_level == "medium"]
merged.text_high <- merged.text[price_level == "high"]

palette_low <- brewer.pal(9, "Blues")[4:9]
palette_mid <- brewer.pal(9, "Purples")[4:9]
palette_high <- brewer.pal(9, "Oranges")[4:9]

tokens_low <- merged.text_low |>
  select(nbd_desc) |>
  unnest_tokens(token, nbd_desc) |>
  filter(!token %in% sw) |>
  filter(!str_detect(token, "[[:digit:]]+")) |>
  count(token, sort = TRUE) |>
  head(20)

tokens_mid <- merged.text_mid |>
  select(nbd_desc) |>
  unnest_tokens(token, nbd_desc) |>
  filter(!token %in% sw) |>
  filter(!str_detect(token, "[[:digit:]]+")) |>
  count(token, sort = TRUE) |>
  head(20)

tokens_high <- merged.text_high |>
  select(nbd_desc) |>
  unnest_tokens(token, nbd_desc) |>
  filter(!token %in% sw) |>
  filter(!str_detect(token, "[[:digit:]]+")) |>
  count(token, sort = TRUE) |>
  head(20)

layout(matrix(1:3, nrow = 1)) 

par(mar = c(2, 2, 2, 2)) 
wordcloud(words = tokens_low$token, 
          freq = tokens_low$n,
          scale = c(2.5, 0.25),  
          max.words = 50,
          colors = palette_low)     
text(x = 0.5, y = -0.2, labels = "Low Price", cex = 1.5, adj = 0.5)

wordcloud(words = tokens_mid$token, 
          freq = tokens_mid$n,
          scale = c(2.5, 0.25),
          max.words = 50,
          colors = palette_mid)
text(x = 0.5, y = -0.2, labels = "Mid Price", cex = 1.5, adj = 0.5)

wordcloud(words = tokens_high$token, 
          freq = tokens_high$n,
          scale = c(2.5, 0.25),
          max.words = 50,
          colors = palette_high)
text(x = 0.5, y = -0.2, labels = "High Price", cex = 1.5, adj = 0.5)

```


**Figure 2: Word Clouds of Neighbourhood Descriptions for Shared Room Listings, by Price Level**

The wordcloud for shared rooms listings differs quite a bit from figure 1 though. This can be for two reasons. The first reason could be that there are not many observations for shared room listings, so perhaps the wordclouds are not representative. Alternatively, the wordclouds to capture some difference in neighbourhood description, compared to the other property types. In that case, it justifies our inclusion of property type as a confounder in predicting price. 

In a similar way, we can also subset for listings in neighbourhoods of a certain crime or income level, to check if the neighbourhood descriptions differ. To do this, we define crime levels and income levels in a similar way to how we defined it for categorical price levels.

However after doing that, we find that there is no substantial difference in wordclouds across the different income and crime levels. This tells us that, at least from the perspective of the listers, the crime or income level does not really impact the neighbourhood description across price levels. 


## Distribution of Price by Property Type

The next visualisation we observe is the stacked histogram of price, by property type. The idea is that we want to verify our previous findings: that is, if price changes by property. 

```{r echo = FALSE, message=FALSE, warning=FALSE}
colors <- c("Entire property" = "#1f77b4",  
            "Private room" = "#ff7f0e",      
            "Shared room" = "#2ca02c")       

p1 <- ggplot(merged, aes(x = price, fill = property_type)) +
  geom_histogram(position = "identity", bins = 30, alpha = 0.5) +
  scale_fill_manual(values = colors) +  
  labs(title = "Stacked Histogram of Price, by Property Type",
       x = "Price", y = "Frequency", fill = "Property Type")

ggplotly(p1)

```


**Figure 3: Stacked Histogram of Price, by Property Type**

As we can see in the above distribution, it does to some degree. The distribution for a private room's price is must narrower and clustered around lower prices, in contrast to entire property. This makes sense, since renting an entire property would make sense to cost a lot more. Most importantly, it justifies our choice to adjust for property type, because as we can see, it does have an impact on our desired response. 

Next we verify whether the price distribution changes according to different levels of scaled median household income, and crime. This can help us inform some of the lack of differences we saw in the wordclouds, when subsetting for different income levels, and crime levels. 

## Boxplots of Price by Neighbourhood Characterstics {.tabset}

### By Neighbourhood Household Income


```{r echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7}

p1 <- ggplot(merged, aes(x = income_level, y = price, fill = income_level)) +
  geom_boxplot() +
  labs(title = "Boxplot of Price by Scaled Median Household Income Level",
       x = "Scaled Income Level",
       y = "Price",
       fill = "Income Level") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggplotly(p1)
```

### By Neighbourhood Crime Level

```{r echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7}

p2 <- ggplot(merged, aes(x = cc_level, y = price, fill = cc_level)) +
  geom_boxplot() +
  labs(title = "Boxplot of Price by Crime Level",
       x = "Crime Level",
       y = "Price",
       fill = "Crime Level") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggplotly(p2)
```

## {-}

**Figure 4: Price Against Scaled Household Income, by Crimes and Property Type**


We note here that the levels for scaled median household income and crime are defined in exactly the same way as they were for price level. 

Analysing the boxplots, there appear to be a substantial amount of outliers for price. However, the actual distribution seems not to really differ between either crime level, or scaled household income. What this suggests, is that even if there is a significant relationship between these variables and price, the actual rate at which price changes with respect to these variables probably will not be that much. 

To verify that claim, we now plot a scatterplot with price as the response. 

## Scatterplot of Price Against Neighbourhood Characteristics, Coloured by Property Type

```{r echo = FALSE, message=FALSE, warning=FALSE}
p <- ggplot(merged, aes(x = price, y = nbd.med_hh_scaled, size = nbd.crime_count, color = property_type)) +
  geom_point(alpha=0.5) +
  scale_size_continuous(range = c(1, 5)) +  
  labs(title = "Price Against Scaled Household Income, by Crimes and Property Type",
       x = "Scaled Household Income (/100)",
       y = "Price",
       size = "Crimes (/Day)",
       color = "Property Type")
p
```

**Figure 5: Price Against Scaled Household Income, by Crimes and Property Type**

Here, we plot price against scaled household income. The size of the dots correspond to increasing crimes levels (that is, crimes committed per day). The colour corresponds to the property type. 

We see that there is at best a weak positive linear association between price and scaled household income. Of course, if there was going to be any association at all, then it makes sense that it would be positive, since if the neighbourhood is affluent, the property is likely to be upscale, and hence cost more. But the fact that it seems that it is not too strong, agrees with our explanation in the introduction of this report. 

In terms of crimes per day, we the majority of the large circles (which mean more crime per day) appear to be clustered around the bottom of the plot - i.e. where price is low. This would give some credence to the claim that more crime leads to lower prices - which is not unreasonable to claim since one would have likely have more crime in less affluent areas. With that said, this does not provide any conclusive evidence towards a relationship, particularly since the small circles (which represent less crimes) appear to be evenly spread across the chart.

Finally, it is worth noting that the majority of the private room listings appear to be in places where household income is lower. On the other hand, those renting out the entire property appear to be evenly spread. There seems to be minimal difference with respect to price though.

## Plotting Distance From Downtown and Neighbourhood Crime and Income, Coloured by Price

```{r message=FALSE, warning=FALSE, fig.height= 7, fig.width=7, echo = FALSE}
sc.plot <- plot_ly(merged, 
                   x = ~nbd.crime_count, 
                   y = ~distance_from_downtown,
                   z = ~nbd.med_hh_scaled,
                   color = ~price, 
                   type = "scatter3d", 
                   mode = "markers",
                   marker = list(size = 8, opacity = 0.5)) |>
  layout(title = "Plot of Neighborhood Crime, Income, and Distance from Downtown, by Price", scene = list(xaxis = list(title = "Neighbourhood Crime Count"),
         yaxis = list(title = "Distance from Downtown"),
         zaxis = list(title = "Neighbourhood Median Household Income")),
         hovermode = "compare")
sc.plot

```

**Figure 6: Scatterplot of Distance to Downtown, Neighbourhood Crime and Income, Coloured by Price**

One variable that we have not yet explored yet is the distance to downtown. We can analyse this alongside the neighbourhood level variables of income level and crime, in a 3D scatterplot, coloured by price. On this visualisation, we see that in general, the majority of the more expensive listings (those of a lighter colour) are found relatively close to downtown, and in neighbourhoods with a slightly lower crime count. Interestingly, the median household income of the neighbourhood does not appear to have any large impact on price, as the proportion of relatively expensive listings (those of a lighter colour) appears to only increase by a little bit, across levels of income. This in particular, agrees with what we found in the previous figures. 

## Map of Neighbourhood Prices Across Toronto

```{r echo = FALSE, message=FALSE, warning=FALSE}
# I take inspiration from my HW1 code here. 

merged.nbd_summary <- merged[, .(
  lon = mean(lon, na.rm = TRUE),
  lat = mean(lat, na.rm = TRUE),
  price = round(mean(price, na.rm = TRUE))
), by = c("nbd")]

downtown_coords <- data.frame(lon = -79.3835, lat = 43.6515)

palette_low <- brewer.pal(9, "Blues")[4:9]

label_html <- lapply(merged.nbd_summary$nbd, function(nbd) {
  HTML(paste("Neighborhood: ", nbd))
})
palette <- colorNumeric(palette = palette_low, domain = merged.nbd_summary$price)

leaflet(merged.nbd_summary) %>%
  addTiles() %>%
  addCircleMarkers(
    ~lon, 
    ~lat,
    color = ~palette(price),
    popup = ~as.character(merged.nbd_summary$nbd),
    label = label_html,
    radius = 8,
    opacity = 0.5, 
    fillOpacity = 1) %>%
  addLabelOnlyMarkers(
    ~lon, ~lat,
    label = ~as.character(merged.nbd_summary$price),  
    labelOptions = labelOptions(
      noHide = TRUE,  
      textOnly = TRUE,  
      direction = "center",
      style = list(color = "white")
    )
  ) %>% 
  addCircleMarkers(
    data = downtown_coords,
    downtown_coords$lon, 
    downtown_coords$lat,
    color = "Purple",
    popup = ~as.character("DT"),
    label = "DT",
    radius = 8,
    opacity = 0.5, 
    fillOpacity = 1) %>%
  addLabelOnlyMarkers(
    downtown_coords$lon, downtown_coords$lat,
    label = ~as.character("DT"),  
    labelOptions = labelOptions(
      noHide = TRUE,  
      textOnly = TRUE,  
      direction = "center",
      style = list(color = "white")
    )
  ) |>
  addLegend('bottomleft', 
            pal = palette, 
            values = merged.nbd_summary$price, 
            title = "Price")

```

**Figure 7: Map of Toronto Neighbourhoods with Mean Price of AirBnB listings**

Our final visualisation is a map where each marker is on a neighbourhood, and written on the marker is the mean price for an AirBnB listing in that neighbourhood. The darker the colour, the higher the price. Hovering over each marker will yield the neighbourhood name. This visualisation is again found on the website. 

We can see that the less costly places to get an AirBnB are on the outskirts of city. And it does appear to be true in general, that as you get closer to downtown (which is marked red on the map), the markers get darker - and hence of higher price. So it does seem to suggest that proximity to downtown does matter towards prices. 


